Application of Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and matrix discrete empirical interpolation

### 2022-07-19 response_to_reviewers.tex

\begin{document}
This statement concerns our revision of the ANUCENE-D-22-00268, entitled ``Modeling Neutronic Transients with Galerkin Projection onto a Greedy-Sampled, POD Subspace'', based on the referees' report.

\Reviewer{\#1}

\begin{response}{
   The paper is clear and relatively well written but has limited novelty. Use of ROM for neutron diffusion problems has already been widely investigated at the Politecnico di Milano (see the work of Stefano Lorenzi and Alberto Sartori), with co-authoring from world-leading experts from SISSA (Professor Rozza), as well as at TAMU (see the work of Peter German and Jean Ragusa). In particular, use of ROM with DEIM for non-linear neutronics is described in:

  "Application of multiphysics model order reduction to doppler/neutronic feedback"
  P German, JC Ragusa, C Fiorina
  EPJ Nuclear Sciences \& Technologies 5 (ARTICLE), 17

@@

  or

  Péter German, Mauricio Tano, Carlo Fiorina, Jean C. Ragusa,
  GeN-ROM—An OpenFOAM®-based multiphysics reduced-order modeling framework for the analysis of Molten Salt Reactors,
  Progress in Nuclear Energy,
  Volume 146,
  2022,
  104148,
  ISSN 0149-1970,
  https://doi.org/10.1016/j.pnucene.2022.104148.

  The novelty of the paper is then restricted to the implementation of the model in a specific code, and to its application to a specific problem. I believe this may still make it worth publishing, but the authors are requested to rewrite abstract, introduction and conclusions in order to:
  - acknowledge the work of other authors on the subject
  - adapt their claims to the fact that the methodology has already been proposed and tested.
  They should also change the title, as it gives the idea that this is a first-of-a-kind paper using a POD-Galerkin and DEIM for neutronics, while this is not the case. The title should focus on elements of actual novelty (implementation in a specific code and application to a benchmark).}

We agree with reviewer that ROM with DEIM was successfully used before and we have acknowledged this in the introduction section and cited these papers.
However, it is important to clarify that in our work, we used a different variant of the DEIM algorithm, which is its matrix version.
Moreover, where DEIM was used previously to approximate non-linear functions such as the temperature and the velocity filed, we have used MDEIM to approximate the problem operator in case an affine decomposition is not readily obtainable. To the author knowledge the matrix version of DEIM was not used before in the nuclear community.
As the reviewer suggested, we have modified the title to reflect this goal
\end{response}

@@

\begin{response}{I would not define JFNK as a modern method. I have found publications from the eighties on it, and I guess it may be older than that. I also would avoid passing the message that block-matrix coupling is faster than operator-splitting. Very often, this is not the case, as operator splitting allows a very well targeted preconditioning, and selective solution of non-converged fields
}

we agree with author. Our intention was that the algorithm is implemented in modern codes, but we have removed the word `modern' to avoid confusion.

@@

 They are sampled from the normal distribution shown in table~4. No correlation was assumed between the six parameters.

\end{response}

\begin{response}
{I find it very difficult to interpret the results of section 4.2.2. The authors compare the predictions of ROM and FOM, but what is the impact of the parameters on the FOM itself? Without this information, it is impossible to assess the goodness of the ROM approximation. The error made by the ROM model should be small with respect to the differences expected in the FOM when changing the selected parameters.}

We mentioned in the paper that the time-to-peak differs from FOM sample to another.
On the other hand, the time-to-peak of the ROM solution is exactly the same as in the FOM solution.
That is why it was not convenient to compute point-by-point difference between the FOM solutions. 
The difference between the FOM peak power at different parameters points varies between 0.1 to 1\%,  which indicates a significantly larger difference between the FOM solution peak powers compared to the ROM error.
we have clarified this in the paper.
------------------------------------
\end{response}

@@

\Reviewer{\#2}
\begin{response}
	{
		Equations (14)-(15), (18)-(20):
		In my understanding, the non-linear operator L is calculated with the MDEIM, and the authors also applied the POD for Equation (14) to treat the temperature distribution with fewer DOFs.
		However, I could not clearly understand how the authors calculate Equation (15) before applying MDEIM from the expansion coefficients $(a_T)$ for the temperature distribution. A few more descriptions of the procedure will help a better understanding for readers.
		( I mean the relationship between the non-linear operator L and f(y(t)) in section 2.2 is not clear. The description of Equation (12) also makes it difficult to clearly understand because Equation (12) can be seen as a linear equation but the actual operator L is a non-linear function due to the feedback effect.)
	}

It is more accurate to say that the operator L is approximated (not calculated) by MDEIM.
For clarity, at each time step, MDEIM is first applied to approximate the operator, and Eq~18 is solved for the flux coefficients.
Theses flux coefficient are then used in Eq.~20 to update the temperature coefficients $a-T$.
Using this computed temperature, the fast cross-section is updated using Eq.~15.
But we agree with reviewer that Eq.~12, as presented, is linear, since was a general representation of the diffusion equation, and this has been clarified in the paper.
In particular, we clarified that in the problem studied here, coupling with TH model leads to $\textbf{A}$ be a function of the solution, i.e, $\textbf{A}(\textbf{y}(t))$.
Also, it is mentioned that vectorized form of the operator corresponds to the function f(y(t)) given in section 2.2.

@@

\end{response}

-------------------------------

\begin{response}
{ pages 15 and Table 1:
	The Doppler feedback coefficient, $3.034x10^-3 K^-0.5$, seems not identical to the benchmark specification in the reference [15]. In the original benchmark specification, it would be $2.034x10^-3 K^-0.5$.
	In addition, the axial bucking $(B^2)$ is specified in the original LRA 2D benchmark problem in my understanding. If the authors are using axial buckling, it should be described in Table 1. "nu" value is also missing in Table 1.
	(You can also find the LRA benchmark results calculated with the conventional codes in the following reference:
	T.M. Sutton, and B. N. Aviles, "Diffusion Theory Methods for Spatial Kinetics Calculations", Prog. Nucl. Energy, 30(2), 119-182, 1996.
	Since the initial keff = 0.9975 is a little higher than their results, I wonder whether the axial buckling is taken into account or not.)
	Off course, the above inconsistency in the FOM calculation condition does not affect the validity of the methodology presented in this work.}

The doppler coefficient .....
The axial buckling was accounted for as given on the benchmark documentation. We have commented on that in the paper and added its value in table~1 along with the value of $\nu$. Thanks for the catch.
Regarding to the value of initial keff, it is sensitive to the spatial discretization, the numerical solver and the convergence criteria, and this might be the cause of this little bias that you observed.
We agree with the reviewer that any inconsistency resulting from a change in the problem data from that in the original benchmark does not affect the presented results, as long the problem parameters are consistent in both the FOM and the ROM.

### 2022-07-19 response_to_reviewers.tex

\usepackage{microtype,xparse,tcolorbox}
\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
      \ttfamily #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

\title{Statement on the Revision of ANUCENE-D-22-00268 \\
  Based on the Referees' Report}
\author{Rabab Elzohery \and Jeremy A. Roberts}
\date{\today}

\begin{document}
This statement concerns our revision of the ANUCENE-D-22-00268, entitled ``Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and DEIM'', based on the referees' report.

### 2022-07-19 paper/main.tex

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%POD-Galerkin projection
%\title{Application of the Matrix version of DEIM and POD-Galerkin projection 
\title{Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and DEIM}
%% \tnotetext[label1]{}
\author[focal]{Rabab Elzohery\corref{cor}}
\ead{Rababelzohery@ksu.edu}
%\ead{http://www.rine.u-fukui.ac.jp}
\author{Jeremy Roberts}
%% \fntext[label2]{}
\cortext[cor1]{Corresponding author}
\address {The Alan Levin Department of Mechanical and Nuclear Engineering,
	Kansas State University \\
	3002 Rathbone Hall, Manhattan, KS 66506, US \\ rababelzohery@ksu.edu, jaroberts@ksu.edu}

@@

%------------------------------------------------------------------------------
% 
%------------------------------------------------------------------------------
\begin{abstract}
Presented here is an application of model-order reduction to non-linear neutronic transients.
The effort described extends previous work\cite{elzohery2021exploring} aimed to develop and assess a Reduced-order model(ROM) framework that can be used for different purposes, such as design optimization and propagating input parameter uncertainties like nuclear data.
Here, the target applications is non-linear.
The ROM was built with the use of the POD-Galerkin projection method and the matrix version of the Discrete Empirical Interpolation Method (MDEIM) for improved efficiency.
Where applications of the ROM to the diffusion equation have been introduced before \cite{}, this work focuses on treating the non-linearity induced by coupling with different physics.
Previous work has used the DEIM to approximate non-linear fielsa such as the temperature \cite{german2019application}.
Here, the MDEIM is used to approximate the projection of the problem operator that otherwise must be computed at each non-linear iteration within each time step, which increases the ROM cost.
By applying the method to the LRA benchmark, speed ups from 6 to 40 were achieved depending on the full-order model (FOM) spatial fidelity.
The maximum relative error in the assemblies power was in the order of $1\times10^{-5}$\%.
Moreover, the ROM was parameterized by POD-greedy sampling, and a case study was performed where the kinetic data, i.e, the precursor decay constants and delayed-neutron fraction were assumed to be the model parameters of interest.
For the parameterized ROM, the maximum relative error in the assembly power was $1\times 10^{-3}$\%, 
The computational gain from the MDEIM-ROM was compared with that of a direct projection ROM and showed that employing the MDEIM is computationally advantageous, and the advantage grows with dimension of the problem.
The method was implemented in the open-source deterministic transport code Detran\cite{roberts2014advanced} where only access to the system operator was needed to construct the ROM, and; hence, the method is trivially invasive and could easily be implemented in many modern deterministic diffusion codes.

@@

A modern algorithm such as Jacobian free Newton-Krylov (JFNK) \cite{gaston2009parallel} can be used to perform this implicit coupling.

@@

In this study, the developed ROM was extended to treat the non-linearity induced by multi-physics modeling.
At this point, we must emphysize that the method used here, i.e; POD Galerkin projection has been used in previous works to approximate the diffusion equation \cite{} for different problems and codes where it showed promising results.
Moreover, the  Discrete Empirical Interpolation Method, DEIM, has been used in conjunction with the POD-Galerkin projection to approximate non-linear fields such as the temperature and the flow fields  .... \cite{}.  

@@

The primary purpose of using the MDEIM is to obtain an offline-online decomposition of the ROM to avoid repetitive projection of the problem operator at each iteration within the time step.

@@

Thus, 50 new parameter samples were generated, and the ROM was executed with the resultant greedy basis.
Moreover, the FOM was run at these testing parameter points, and the error between both models was computed.
The quantities of interest were the assembly power at the peak and the maximum core temperature, i.e., the end-of-transient temperature.
Note that the time to peak is different from sample to sample.
Shown in Fig.~ \ref{fig:peak power mean error}, the sample mean of the assembly power error, at the peak, where the mean core maximum temperature error is shown in Fig.~\ref{fig:T mean error}.

>>>>

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%POD-Galerkin projection
\title{Application of Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and matrix discrete empirical interpolation}
%\title{Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and DEIM}
%% \tnotetext[label1]{}
\author[focal]{Rabab Elzohery\corref{cor}}
\ead{Rababelzohery@ksu.edu}
%\ead{http://www.rine.u-fukui.ac.jp}
\author{Jeremy Roberts}
%% \fntext[label2]{}
\cortext[cor1]{Corresponding author}
\address {The Alan Levin Department of Mechanical and Nuclear Engineering,
	Kansas State University \\
	3002 Rathbone Hall, Manhattan, KS 66506, US \\ rababelzohery@ksu.edu, jaroberts@ksu.edu}

@@

%------------------------------------------------------------------------------
% 
%------------------------------------------------------------------------------
\begin{abstract}
Presented here is an application of model-order reduction to non-linear neutronic transients.
The effort described extends previous work\cite{elzohery2021exploring} aimed to develop and assess a Reduced-order model (ROM) framework that can be used for different purposes, such as design optimization and propagating input parameter uncertainties like nuclear data.
Here, the target applications is non-linear.
The ROM was built with the use of the POD-Galerkin projection method and the matrix version of the Discrete Empirical Interpolation Method (MDEIM) for improved efficiency.
Where applications of the ROM to the diffusion equation have been introduced before \cite{sartori2014comparison, sartori2014reduced, tsujita2021fast}, this work focuses on treating the non-linearity induced by coupling with different physics.
Previous work has used DEIM to approximate non-linear functions resulting from multiphysics coupling such as the temperature and the velocity field \cite{german2019application, german2022gen}.
Here, the MDEIM is used to approximate the projection of the problem operator that otherwise must be computed at each non-linear iteration within each time step, which increases the ROM cost.
By applying the method to the LRA benchmark, speed ups from 6 to 40 were achieved depending on the full-order model (FOM) spatial fidelity.
The maximum relative error in the assemblies power was in the order of $1\times10^{-5}$\%.
Moreover, the ROM was parameterized by POD-greedy sampling, and a case study was performed where the kinetic data, i.e, the precursor decay constants and delayed-neutron fraction were assumed to be the model parameters of interest.
For the parameterized ROM, the maximum relative error in the assembly power was $1\times 10^{-3}$\%, 
The computational gain from the MDEIM-ROM was compared with that of a direct projection ROM and showed that employing the MDEIM is computationally advantageous, and the advantage grows with dimension of the problem.
The method was implemented in the open-source deterministic transport code Detran\cite{roberts2014advanced} where only access to the system operator was needed to construct the ROM, and; hence, the method is trivially invasive and could easily be implemented in many modern deterministic diffusion codes.

@@

An algorithm such as Jacobian free Newton-Krylov (JFNK) \cite{gaston2009parallel} can be used to perform this implicit coupling.

@@

In this study, the developed ROM was extended to treat the non-linearity induced by multi-physics modeling, using the matrix version of the  Discrete Empirical Interpolation Method (DEIM).
To this point, we must emphasize that the method used here, i.e; POD Galerkin projection has been used in previous works to approximate the diffusion equation \cite{sartori2014comparison, sartori2014reduced} for other problems using different codes where it showed promising results.
Moreover, DEIM has been used in conjunction with the POD-Galerkin projection to approximate non-linear fields such as the temperature and the velocity field \cite{german2019application, german2022gen}.  

@@

The primary purpose of using the MDEIM is to obtain an offline-online decomposition of the ROM to avoid repetitive projection of the problem operator at each iteration within the time step, in cases that an affine decomposition of the operator is not readily obtainable.

@@

Thus, 50 new parameter samples were generated, and the ROM was executed with the resultant greedy basis.
Moreover, the FOM was run at these testing parameter points, and the error between both models was computed.
The quantities of interest were the assembly power at the peak and the maximum core temperature, i.e., the end-of-transient temperature.
Note that the time to peak is different from sample to sample, with the relative difference between the FOM peak power at the parameter vary between 0.1 to 1\%.
Shown in Fig.~ \ref{fig:peak power mean error}, the sample mean of the assembly power error, at the peak, where the mean core maximum temperature error is shown in Fig.~\ref{fig:T mean error}.
